---
title: "2025-CAH"
author: "Josephine McKelvy"
date: "2025-06-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file includes steps to clean and analyze fictional experimental data as part of a hiring exercise for the research analyst/behavioral researcher role at Duke University's Center for Advanced Hindsight (CAH).

# Getting Started

```{r}
packages <- c("DataExplorer",
              "dplyr",
              "ggmosaic",
              "ggplot2",
              "knitr",
              "readr",
              "sjPlot",
              "stats",
              "summarytools",
              "tinytex",
              "vcd",
              "waffle")

# Install packages that are not already installed:
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Load libraries:
lapply(packages, library, character.only = TRUE)
```

# Merging, Cleaning, Transforming Data

```{r}
# Set working directory to load and save files:
setwd("~/2025-CAH")

# Load data sets:
Data_set_A <- readr::read_csv("Data set A.csv")
Data_set_B <- readr::read_csv("Data set B.csv")

# Review the data sets (AKA summarize the dataframes):
summarytools::dfSummary(Data_set_A) 
#There are 1002 duplicates and 1001 distinct identifiers (thus up to 1001 real cases, ranging from 1000 to 2000 as IDs) making up the 2004 observations; not sure what that remaining blank row is.

summarytools::dfSummary(Data_set_B) 
#There are 751 duplicates and 752 distinct identifiers (thus up to 752 real cases, ranging from 1457 to 2381 as IDs) making up the 1504 observations; not sure what that remaining blank row is.

# Merge data frames by ID variable, then clean that data set to reduce duplication of efforts:
contiguous <- merge(Data_set_A, Data_set_B, by="identifier")
summarytools::dfSummary(contiguous) 

# There are multiple duplicates and 371 distinct identifiers common to both data sets (thus up to 371 cases with conditions and outcomes, ranging from 1457 to 2000 as IDs) for the analytic sample. 

# Remove empty rows and duplicate cases. Then summarize this dataframe again:
contiguous <- dplyr::distinct(contiguous) 

# There are still 3 rows with missing conditions.
# Keep cases where condition is not missing:
contiguous <- filter(contiguous,!is.na(condition)) 

# There are 3 character (or raw text) columns (that should be factor variables) and an identifier that is a double (or real number) data type.
# Save categorical variables as factor with new names for the categories:
contiguous <- contiguous %>%
  mutate(
    condition = factor(condition,
                       levels = c("recommendation","control")), #reorder factors for the contingency table
                      #labels = c("recommendation","informational"),
    income = factor(income_level,
                          levels = c("non-LMI","LMI")),
    outcome = factor(increased_contribution,
                                    levels = c("1","closed"), #reorder factors for the contingency table
                                    labels = c("contributed","closed")) 
  )

# Keep the renamed variables:
contiguous <-  contiguous %>%
  select(identifier, condition, income, outcome)
```

# Exploratory Data Analysis

-   <https://bookdown.org/lyzhang10/lzhang_r_tips_book/preface.html>
-   <https://geanders.github.io/RProgrammingForResearch/exploring-data-1.html>
-   <https://cran.r-project.org/web/packages/ggmosaic/vignettes/ggmosaic.html>

```{r}
contiguousDE <- dummify(contiguous)
DataExplorer::create_report(contiguousDE) # creates report.html output
# weak correlations between contribution, condition, and income level
```

## Stacked Bar Chart of condition
- https://www.cedricscherer.com/2021/07/05/a-quick-how-to-on-labelling-bar-graphs-in-ggplot2/#dataviz

```{r}
# Create the values to graph:
condition_ct  <- table(contiguous$condition) #counts for each condition
condition_pct <- prop.table(condition_ct)    #percentages for each condition
condition_tab <- data.frame(condition_ct,condition_pct)
condition_tab <- condition_tab %>%
  rename(condition = Var1, count = Freq, percent = Freq.1) %>%
  select(condition, count, percent) %>%
  arrange(percent)

# Plot:
bar_plot <- ggplot(condition_tab, aes(x = percent, y = condition)) %>% +
  geom_col(fill = "#156082", width = .75, show.legend = FALSE) +
  geom_text(aes(label = paste(round(percent*100),"%")), position = position_dodge(width = 1), hjust = -0.5) +
  theme_void() +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 14, hjust = 1)) +
  labs(title = "Figure 1. Stacked bar chart of conditions") 

# Save plot:
ggsave(bar_plot, filename = "plot-bar.png", height = 4, width = 12)
```

## Donut chart of income
(not sure why geom_rect didn't work but geom_col did):
- https://r-charts.com/part-whole/donut-chart-ggplot2/
- https://rfortherestofus.com/2022/09/how-to-make-a-donut-chart-in-ggplot

```{r}
# Create the values to graph:
# https://r-graph-gallery.com/128-ring-or-donut-plot.html
income_ct  <-  table(contiguous$income) #counts for each income level
income_pct <-  prop.table(income_ct)    #percentages for each income level
income_ymax <- cumsum(income_pct)       #cumulative percentages of each income level
income_ymin <- c(0, head(income_ymax, n = -1))
  
donut_tab <- data.frame(income_ct, income_pct,income_ymax,income_ymin)
donut_tab <- donut_tab %>%
  rename(income = Var1, count = Freq, percent = Freq.1) %>%
  select(income, count, percent, income_ymax, income_ymin)

# Plot:
hsize <- 4 # 1=small hole size & thick donut, 10=large hole size & thin donut

donut_tab <- donut_tab %>% 
  mutate(x = hsize)

donut_plot <- ggplot(donut_tab, aes(x = hsize, y = percent, fill = income)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste(round(percent*100), "% ", income)), position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y",      # A donut chart is a bar chart with polar coordinates. 
              direction = -1) + # Set the direction to -1 so the filled in part starts at the top and goes clockwise.
  xlim(c(0.2, hsize + 0.5)) +
  scale_fill_manual(values = c("LMI"="#156082","non-LMI"="#E7EAF3")) +
  theme_void() # Removes grid lines and background +
  labs(title = "Figure 2. Donut chart of income levels")

# Save plot:
ggsave(donut_plot, filename = "plot-donut.png")
```

## Waffle Chart of action
- https://r-charts.com/part-whole/waffle-chart-ggplot2/#geom-waffle

```{r}
# Create the values to graph:
waffle_ct  <-  table(contiguous$outcome)   #counts for each outcome
waffle_pct <-  prop.table(waffle_ct) * 100 #percentages for each outcome to standardize to a 10x10 grid plot

# Plot:
waffle_plot <- waffle(waffle_pct, rows = 10,
       colors = c("contributed"="#156082","closed"="#E7EAF3"),
       legend_pos = "none")

# Save plot:
ggsave(waffle_plot, filename = "plot-waffle.png") 
```

## Mosaic Plot of the action by condition and income
- https://rstudio-pubs-static.s3.amazonaws.com/584765_5ab02919bd374db7ad7c58f20a11e86f.html
- https://cran.r-project.org/web/packages/ggmosaic/vignettes/ggmosaic.html

```{r}
# Optional: Create values for labels:
mosaic_ct <- table(contiguous$income, contiguous$condition, contiguous$outcome)
mosaic_pct <- prop.table(mosaic_ct)
mosaic_tab <- data.frame(mosaic_ct,mosaic_pct)
mosaic_tab <- mosaic_tab %>%
  rename(income = Var1, 
         condition = Var2, 
         outcome = Var3, 
         conditionXincome = Freq) %>%
  select(income, condition, outcome, conditionXincome)

# Create a mosaic plot (or percent stacked bar chart) of the contingency table:
ggsave(filename = "plot-mosaic_1.png",
       mosaic_1 <- ggplot(data = contiguous) +
         ggmosaic::geom_mosaic(aes(x = product(outcome), fill = income), # normally, fill is your outcome variable
                  divider = c("vspine","hbar"), # vspine keeps the column widths constant & hbar lets the heights vary
                  offset = 0.02) +
         facet_grid(~condition) +
         scale_fill_manual(values = c("LMI"="#156082","non-LMI"="#E7EAF3")) +
         #theme_void() +
         theme(panel.background = element_blank(),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank()) +
         labs(title = "Figure 4. Mosaic plot of income and increased contributions, by condition"))

# Almost all workers increased contributions, regardless of condition.
# More non-LMI workers received the recommendation email.
# More LMI workers received the informational (control) email.

# Create and save a mosaic plot with Pearson residuals:
jpeg(filename = "plot-Mosaic_2.png")
mosaic_2 <- vcd::mosaic(~ outcome + condition + income, 
                        data = contiguous, 
                        main = "Retirement contributions by condition and income", 
                        shade = TRUE, legend = TRUE)
# Statistically (but not substantively) significant correlations
```

# Statistical Analysis

-   Null Hypothesis: Treated participants (receiving a recommendation email that leverages peer information) are just as likely as control participants (receiving a generic informational email) to increase their retirement contributions.

```{r}
# Create cross-tabs of the frequencies/counts for the categorical outcome by condition:
ctab <- table(contiguous$condition, contiguous$outcome)
summary.table(ctab) # Chisq = 0.08518, df = 1, p-value = 0.7704
# At least 2 cells have frequencies that are less than 5, so a Fisher's exact test may be more appropriate than chi-square. 
stats::fisher.test(ctab) # p-value = 1
# Workers who received the recommendation email were 34% more likely (than those who received the informational email) to increase their TSP contribution (OR = 1.338,95% CI 0.09-18.65), if these results were statistically significant. Overall, just about everyone increased contributions, regardless of condition.

# Export the cross-tab: 
sjPlot::sjt.xtab(contiguous$outcome, contiguous$condition,
                 title = "Table 1. Contingency table of contributions by email condition",
                 file = "table-condition.doc")

# Explore other variables:
itab <- table(contiguous$income, contiguous$outcome)
summary.table(itab) # Chisq = 3.936, df = 1, p-value = 0.04727
# statistically significant but at least 2 cells have frequencies that are less than 5.
stats::fisher.test(itab) # p-value = 0.12
# Workers who closed (their accounts?) were categorized as low-to-moderate income.

# Export the cross-tab:
sjPlot::sjt.xtab(contiguous$outcome, contiguous$income,
                 title = "Table 2. Contingency table of contributions by income level",
                 file = "table-income.doc")
```

# Conclusion

-   On average, receiving an email prompted federal workers to increase their retirement contributions. Those who closed (their accounts?) were all employees with low-to-moderate income. There was no relationship between the type of email that employees received and their subsequent action.
